Toward single particle reconstruction without particle picking: Breaking the detection limit

### 2018-09-07 pnas/Towards.tex
Responding to Amit's comments

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a]{Tamir Bendory}
\author[b]{Nicolas Boumal} 
\author[c]{William Leeb}
\author[a]{Eitan Levin}
\author[a,b]{Amit Singer}

@@

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{Please provide details of author contributions here.}
\authordeclaration{Please declare any conflict of interest here.}
%\equalauthors{\textsuperscript{1}A.O.(Author One) and A.T. (Author Two) contributed equally to this work (remove if not applicable).}
\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: author.two\@email.com}

@@

\dropcap{C}ryo--electron microscopy (cryo--EM) is an leading technology for single particle reconstruction (SPR) of macromolecules. \TODO{revise first sentence}
In a cryo--EM experiment, biological samples are rapidly frozen in a thin layer of vitreous ice. %Within the ice, the molecules are randomly oriented and positioned.
 The microscope produces a 2-D tomographic image of the samples embedded in the ice, called a \emph{micrograph}. Each micrograph contains tomographic projections of the samples at unknown locations and under unknown viewing directions. The goal is to construct 3-D models of the molecules from the micrographs.

The signal to noise ratio (SNR) of the projections in the micrographs is a function of two dominating factors. On the one hand, the SNR is a function of the electron dose. To keep radiation damage within acceptable bounds, the dose must be kept low, which leads to high noise levels. On the other hand, the SNR is a function of the molecule size. The smaller the molecules, the fewer detected electrons carry information about them. \TODO{Amit: do you agree with the last sentence? maybe we should say it is a function of the contrast?}

All contemporary methods in the field split the reconstruction procedure into several stages.
The first stage consists in extracting the various particle projections from the micrographs. This stage is called \emph{particle picking}. Later stages aim to construct a 3-D model of the molecule from these projections. The quality of the reconstruction eventually hinges on the quality of the particle picking stage. Figure~\ref{fig:micro_example} illustrates how particle picking becomes increasingly challenging as the SNR degrades.

@@

Crucially, it can be shown that reliable detection of individual particles is impossible below a certain critical SNR. This fact has been recognized early on by the cryo-EM community. In particular, in an influential paper from 1995, Richard Henderson~\cite{henderson1995limitations} investigates the following questions:
\begin{myquote}{0.15in}
	\emph{For the purposes of this review, I would like to ask the question: what is the smallest size of free-standing molecule whose structure can in principle be determined by phase-contrast electron microscopy? Given what has already been demonstrated in published work, this reduces to the question: what is the smallest size of molecule for which it is possible to determine from images of unstained molecules the five parameters needed to define accurately its orientation (three parameters) and position (two parameters) so that averaging can be performed?}
\end{myquote}
%\begin{quote}
%	\emph{For the purposes of this review, I would like to ask the question: what is the smallest size of free-standing molecule whose structure can in principle be determined by phase-contrast electron microscopy? Given what has already been demonstrated in published work, this reduces to the question: what is the smallest size of molecule for which it is possible to determine from images of unstained molecules the five parameters needed to define accurately its orientation (three parameters) and position (two parameters) so that averaging can be performed?}
%\end{quote}
In that paper and in others that followed (e.g.,~\cite{glaeser1999electron}), it was established that particle picking is impossible for molecules below a certain weight (below $\sim$50 kDa). 
Joachim Frank voices a similar observation in his 2017 Nobel prize lecture: ``\emph{Using the ribosome as an example, it became clear from the formula we obtained that the single-particle approach to structure research was indeed feasible for molecules of sufficient size: Particle Size > 3/[Contrast$^2$ $\times$ Resolution (as length) $\times$ Critical Electron Dose]}''~\cite{frank2018single}. 
\TODO{As these two leaders of the cryo-EM community point out,} it is impossible to reconstruct small molecules by any of the existing computational pipelines for single particle analysis in cryo--EM, because the particles themselves cannot be picked from the micrographs.

The unique issues raised by small particles have been mitigated by recent technological advances in the field, including the use of Volta phase plates~\cite{khoshouei2017cryo,liang2017phase} and scaffolding cages~\cite{liu2018nearatomic}.
Despite this progress, detecting small molecules in the micrographs remains a challenge.
We note that nuclear magnetic resonance (NMR) spectroscopy and X-ray crystallography are well suited to reconstruct small molecules. Yet, cryo--EM has a lot to offer even for molecules with already known structures obtained via NMR spectroscopy or X-ray crystallography, because these methods have limited ability to distinguish conformational variability. \TODO{Need a ref for this claim.}

@@

As a proof of concept, we first study a toy model \TODO{Amit: do you want to keep ``toy model'' or replace with ``mathematical toy model'', ``simplified model'' etc.} for which it is easier to convey the mathematical principles. 
In this toy model, an unknown image appears multiple times at unknown locations in each of several micrographs, each affected by additive Gaussian noise---see Figure~\ref{fig:micro_example} for an illustration.
The goal is to estimate the planted image. The task is challenging in particular when the SNR is low enough that particle picking (identifying the locations of each image in each micrograph) cannot be done reliably. 
This problem is interesting of its own as it appears in other scientific applications, including spike sorting~\cite{lewicki1998review}, passive radar~\cite{gogineni2017passive} and system identification~\cite{ljung1998system}.

In order to recover the image, we use autocorrelation analysis. Specifically, we relate the autocorrelations of the micrographs to the autocorrelations of the image.
For any noise level, these autocorrelations can be estimated to any desired accuracy, provided that we  observe sufficiently many image occurrences and the latter are separated in the micrograph. Importantly, there is no need to detect individual image occurrences. The autocorrelations of the micrographs are straightforward to compute and require only one pass over the data. After estimation of the density of particles in the micrographs, these directly yield estimates for the autocorrelations of the target image itself. To estimate the image itself from its estimated autocorrelations, we solve a nonlinear inverse problem.

Beyond this 2-D proof of concept, we look toward \mbox{3-D} reconstruction as well. Zvi Kam~\cite{kam1980reconstruction} first proposed autocorrelation analysis for \mbox{3-D} reconstruction, under the assumption of perfect particle picking: his method used autocorrelations of the picked, perfectly centered, particles. In contrast, we derive the mathematical relation between the autocorrelations of the micrographs as a whole and the \mbox{3-D} volume, under some simplifying conditions.
We show a few numerical examples and outline the future developments required to make this method applicable on real data.

@@

\TODO{Need to cite \cite{scapin2018cryo}. Discusses cryo-em for drug discovery and mentions small molecule}

@@

Let $x\in\RL$ be the target signal and let $y\in\RN$ be the observed data, where $N$ is assumed far larger than $L$. Let  $s \in \{0, 1\}^{N-L+1}$ be a binary signal indicating (with 1's) the starting positions of all occurrences of $x$ in $y$, so that, with additive white Gaussian noise:
\begin{align}
y & =  x \ast s + \varepsilon, & \varepsilon & \sim \mathcal{N}(0,\sigma^2 I_N),
\label{eq:model}
\end{align}
where $\ast$ denotes linear convolution. 
While both $x$ and $s$ are unknown, the goal is only to estimate $x$ from $y$.
This is a key difference with other works on  \emph{blind deconvolution}, a longstanding problem arising in a variety of engineering and scientific applications such as astronomy, communication, image deblurring, system identification and optics; see~\cite{jefferies1993restoration,shalvi1990new,ayers1988iterative,abed1997blind}, just to name a few. 
The parameters of the signal $s$ (the locations of its nonzero values) are the \emph{nuisance variables} of the problem. As will be shown next, in a low SNR environment, estimating $s$ is impossible, whereas estimating $x$ is tractable under some conditions on $s$.

@@

This decision problem can be abstracted as follows: We have two known vectors $\theta_0 = x$ and $\theta_1 = 0$ in $\mathbb{R}^L$. There is a random variable $\eta$ taking values 0 or 1 with probabilities $q$ and $1-q$, respectively. We observe a random vector $X \in \mathbb{R}^L$ (an extract of the micrograph) with the following distribution: if $\eta = 0$, then $X \sim \mathcal N(\theta_0,\sigma^2)$ and if $\eta = 1$, then $X \sim \mathcal N(\theta_1,\sigma^2)$.

@@

In general, for a signal $z$ of length $m$, the autocorrelation of order $q = 1, 2, \ldots$ is given for any integer shifts $\ell_1, \ldots, \ell_{q-1}$ by
\begin{align}
a_z^q[\ell_1,\ldots,\ell_{q-1}]  & = \frac{1}{m} \sum_{i=-\infty}^{+\infty} z[i]z[i+\ell_1]\cdots z[i+\ell_{q-1}],
\label{eq:ac_general}
\end{align}
where indexing of $z$ out of the range $0, \ldots, m-1$ is zero-padded.
For our purposes, this will be applied both to $x$ (of length $L$) and to $y$ (of length $N$).
Explicitly, the first-, second- and third-order autocorrelations are given by: \TODO{we may want to remove these explicit formulas}
\begin{align} 
a_z^1 & = \frac{1}{m} \sum_{i=0}^{m-1} z[i], \nonumber\\
a_z^2[\ell] & = \frac{1}{m} \sum_{i = \max\{0, -\ell\}}^{m-1 + \min\{0, -\ell\}} z[i]z[i+\ell], \nonumber\\
a_z^3[\ell_1,\ell_2] & = \frac{1}{m} \sum_{i = \max\{0, -\ell_1, -\ell_2\}}^{m-1 + \min\{0, -\ell_1, -\ell_2\}} z[i]z[i+\ell_1]z[i+\ell_2]. \label{eq:ac_special}
\end{align}
The autocorrelation functions have symmetries. Specifically, $a_z^2[\ell] = a_z^2[-\ell]$, and
$a_z^3[\ell_1,\ell_2] = a_z^3[\ell_2,\ell_1]=a_z^3[-\ell_1,\ell_2-\ell_1].
$

@@

For shifts in $0, \ldots, L-1$, the autocorrelation functions of $y$ depend on the corresponding autocorrelations of $x$, the noise level $\sigma$ and the support signal $s$. Importantly, under the separation condition~\eqref{eq:spacing}, the dependency on $s$ is only through the density $\gamma$.
We consider the asymptotic regime where $M, N\to\infty$, while $\gamma$ remains constant (we see an increasingly large micrograph, containing increasingly many signal occurrences, with constant signal density). In that regime, the law of large numbers can be used to show the following statement:
\begin{align} \label{eq:mean_micrograph}
\lim_{N\to\infty} a_y^1 & \aseq \gamma a_{x}^1,
\end{align}
where equality holds almost surely (a.s.), meaning it holds with probability one.
The randomness is over the Gaussian noise $\varepsilon$; $s$ may be deterministic.
Thus, given enough data, if $\gamma$ is known, we can estimate $a_x^1$ from $y$. (We show later how to estimate $\gamma$ as well.)

We have a similar observation for the second-order autocorrelation: $a_y^2[\ell]$ computes the correlation between $y$ and a copy of $y$ shifted by $\ell$ entries. Considering $\ell$ only in the range $0, \ldots, L-1$, one can see that any given occurrence of $x$ in $y$ is only ever correlated with itself, and never with another occurrence. As a result,
\begin{align}
	\lim_{N\to\infty} a_y^2[\ell] & \aseq \gamma a_{x}^2[\ell] + \sigma^2\delta[\ell],
	\label{eq:ac2_micrograph}
\end{align}
for $\ell = 0, \ldots, L-1$, where $\delta$ denotes the Kronecker delta function. The last part captures the autocorrelation of the noise. Notice that, even if $\sigma$ is unknown, entries $\ell = 1, \ldots, L-1$ still provide useful information about $a_x^2$.

Along the same lines, one can establish a relation for third-order autocorrelations:
\begin{multline}
\lim_{N\to\infty} a_y^3[\ell_1,\ell_2] \aseq \gamma a_{x}^3[\ell_1,\ell_2] \label{eq:ac3_micrograph} \\ + \sigma^2\gamma a_{x}^1 \cdot \big(\delta[\ell_1,0]+\delta[0,\ell_2]+\delta[\ell_1,\ell_2]\big),
\end{multline}
for $\ell_1,\ell_2 = 0, \ldots, L-1$. Here too, few entries are affected by $\sigma$ in the limit.
Detailed derivations for identities in this and the next part are given in Appendix~\ref{sec:autocorrelation_computation}.

@@

Some remarks are in order. First, \eqref{eq-uniqueness} is not numerically stable: if $z[0]$ or $z[L-1]$ are close to 0, recovery of $z$ is sensitive to errors in the autocorrelations. In practice, we recover $z$ by fitting it to its autocorrelations using a nonconvex least-squares (LS) procedure, which is empirically more robust to additive noise; we have observed similar phenomena for related problems~\cite{bendory2017bispectrum,boumal2017heterogeneous,abbe2017multireference}.
Second, if the separation condition holds, then the length of the signal can be determined from the autocorrelations.
In particular, if  $a_y^2[L-i] = 0$ for $i=1,\ldots,T$,
then the support of $x$ has length $L-T$ rather than $L$. 
Third, note that the second-order autocorrelation is not by itself sufficient to determine the signal uniquely. %~\cite{beinert2015ambiguities,bendory2017fourier}.
However, for dimensions greater than one, almost all signals are determined uniquely up to sign (or phase for complex signals) and reflection through the origin (with conjugation in the complex case). %~\cite{hayes1982reconstruction,hayes1982reducible}. 
The sign ambiguity can be resolved by the mean of the signal if it is not zero. However, determining the reflection symmetry still requires additional information, beyond the second-order autocorrelation.

The observed moments $a_y^1,a_y^2$ and $a_y^3$ of $y$ do not immediately yield the moments of the signal $x$, as seen by~\eqref{eq:mean_micrograph},~\eqref{eq:ac2_micrograph} and~\eqref{eq:ac3_micrograph}; rather, the two are related by the noise level $\sigma$ and the ratio $\gamma$. We will show, however, that $x$ is still identifiable from the observed moments of $y$. In general, we say a parameter is ``identifiable'' if its value is uniquely determined in the limit $N \to \infty$.

First, we observe that if the noise level $\sigma$ is known, generally, one can estimate $\gamma$ from the first two moments of the micrograph.
%
\begin{proposition} \label{prop:gamma}
	Let $\sigma > 0$ be fixed. If the mean of $x$ is nonzero, then 
	%
	\begin{align*}
	%
	\gamma & \aseq \lim_{N \to \infty}\frac{L (a^1_y)^2}{a_y^2[0] + 2\sum_{\ell = 1}^{L-1}a_y^2[\ell]-\sigma^2}.
	%
	\end{align*}
	%
\end{proposition}
\begin{proof}
In the limit, $(a^1_y)^2=\frac{\gamma^2}{L^2}\sum_{i=0}^{L-1}\sum_{j=0}^{L-1}x[i]x[j]$. Similarly,  $\sum_{\ell = 1}^{L-1}a_y^2[\ell]=\frac{\gamma}{L}\sum_{\ell = 1}^{L-1}\sum_{i=0}^{L-1-\ell}x[i]x[i+\ell]$ and $a_y^2[0]=\frac{\gamma}{L}\sum_{i=0}^{L-1}x^2[i] + \sigma^2$. The proof is concluded by noting that  $a_x^2[-\ell]=a_x^2[\ell]$. 
%	Plug in the explicit expressions of~\eqref{eq:mean_micrograph} and~\eqref{eq:ac2_micrograph} into the right hand side of the equality.
\end{proof}

Using third-order autocorrelation information of $y$, both the ratio $\gamma$ and the noise $\sigma$ are identifiable. For the following results, when we say that a result holds for a ``generic'' signal $x$, we mean that the set of signals which cannot be determined by these measurements
lies in a set with Lebesgue measure zero. 
In particular, this means that we can recover
almost all signals with the given measurements.

\begin{proposition} \label{prop:gamma_sigma}
	%
	%Let $\sigma > 0$ be fixed. 
	Assume $L \geq 3$. % Added by Nicolas, Sep 4, 2018: I checked all identities for L >= 3, with some computer code; I didn't write code to verify the corner cases L = 1 and L = 2.
	The observed autocorrelations $a_y^1,a_y^2$ and  $a_y^3$ determine the ratio $\gamma$ and noise level $\sigma$ uniquely for a generic signal $x$. If $\gamma > \frac{1}{4}$, then this holds for any signal $x$ with nonzero mean. 
	\begin{proof}
		See Appendix~\ref{sec:proof_prop_gamma_sigma}.
	\end{proof}
\end{proposition}

From Propositions~\ref{prop:uniqueness} and~\ref{prop:gamma_sigma} we can directly deduce the following:
\begin{corollary}
	%Let $K=1$ and $\sigma > 0$ be fixed.
	 The signal $x$, the ratio $\gamma$, and the noise level $\sigma$ are identifiable from the first three autocorrelation functions of $y$ if:
	\begin{itemize}
		\item Either the signal $x$ is generic; or
		\item   $x$ has nonzero mean  and $\gamma > \frac{1}{4}$.
	\end{itemize}
\end{corollary}
%
%\TODO{Similar results hold for signals in higher dimensions.---need to keep this sentence in red until the very end to make sure it's correct.}

@@

@@

>>>>

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a]{Tamir Bendory}
\author[b]{Nicolas Boumal} 
\author[c]{William Leeb}
\author[a,b]{Eitan Levin}
\author[a,b]{Amit Singer}

@@

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{Please provide details of author contributions here. \revised{Let's leave it open for now} }
\authordeclaration{The authors declare no conflicts of interest.}
%\equalauthors{\textsuperscript{1}A.O.(Author One) and A.T. (Author Two) contributed equally to this work (remove if not applicable).}
\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: author.two\@email.com}

@@

\dropcap{C}ryo--electron microscopy (cryo--EM) is an imaging technique in structural biology used for single particle reconstruction (SPR) of macromolecules.
In a cryo--EM experiment, biological samples are rapidly frozen in a thin layer of vitreous ice. %Within the ice, the molecules are randomly oriented and positioned.
 The microscope produces a 2-D tomographic image of the samples embedded in the ice, called a \emph{micrograph}. Each micrograph contains tomographic projections of the samples at unknown locations and under unknown viewing directions. The goal is to construct 3-D models of the molecular structure from the micrographs.

The signal to noise ratio (SNR) of the projections in the micrographs is a function of two dominating factors. On the one hand, the SNR is a function of the electron dose. To keep radiation damage within acceptable bounds, the dose must be kept low, which leads to high noise levels. On the other hand, the SNR is a function of the molecule size. The smaller the molecules, the fewer detected electrons carry information about them.

All contemporary methods in the field split the reconstruction procedure into two main  stages.
The first stage consists in extracting the various particle projections from the micrographs. This stage is called \emph{particle picking}. The second stage aims to construct a 3-D model of the molecular structure from these projections. The quality of the reconstruction eventually hinges on the quality of the particle picking stage. Figure~\ref{fig:micro_example} illustrates how particle picking becomes increasingly challenging as the SNR degrades.

@@

\begin{figure*}[h]
	\centering
	\begin{subfigure}[h]{0.24\textwidth}
		\centering
		\includegraphics[scale=0.35]{Einstien_progress_1}
		\caption{$500$ micrographs \TODO{Tamir: I am redo the exp}}
	\end{subfigure} 
	\begin{subfigure}[h]{0.24\textwidth}
		\centering
		\includegraphics[scale=0.35]{Einstien_progress_4}
		\caption{$2\times 10^3$ micrographs}
	\end{subfigure} 
	\begin{subfigure}[h]{0.24\textwidth}
		\centering
		\includegraphics[scale=0.35]{Einstien_progress_40}
		\caption{$2\times 10^4$ micrographs}
	\end{subfigure} 
	\begin{subfigure}[h]{0.24\textwidth}
		\centering
		\includegraphics[scale=0.35]{Einstien_progress_400}
		\caption{$2\times 10^5$ micrographs}
	\end{subfigure}
	\caption{\label{fig:Einst_example} Recovery of Einstein from micrographs at noise level $\sigma = 3$ (see Figure~\ref{fig:micro_example}(c)). Averaged autocorrelations of the micrographs allow to estimate the power spectrum of the target image. This does not require particle picking. A phase retrieval algorithm (RRR) produces the estimates here shown, initialized randomly. Estimates are obtained from $500,2\times 10^3,2\times 10^4,2\times 10^5$ micrographs (growing across panels), each containing $700$ image occurrences on average.}	
\end{figure*}

Crucially, it can be shown that reliable detection of individual particles is impossible below a certain critical SNR. This fact has been recognized early on by the cryo-EM community. In particular, in an influential paper from 1995, Richard Henderson~\cite{henderson1995limitations} investigates the following questions:
\begin{myquote}{0.15in}
	\emph{For the purposes of this review, I would like to ask the question: what is the smallest size of free-standing molecule whose structure can in principle be determined by phase-contrast electron microscopy? Given what has already been demonstrated in published work, this reduces to the question: what is the smallest size of molecule for which it is possible to determine from images of unstained molecules the five parameters needed to define accurately its orientation (three parameters) and position (two parameters) so that averaging can be performed?}
\end{myquote}
In that paper and in others that followed (e.g.,~\cite{glaeser1999electron}), it was established that particle picking is impossible for molecules below a certain weight (below $\sim$50 kDa). 
Joachim Frank voices a similar observation in his 2017 Nobel prize lecture: ``\emph{Using the ribosome as an example, it became clear from the formula we obtained that the single-particle approach to structure research was indeed feasible for molecules of sufficient size: Particle Size > 3/[Contrast$^2$ $\times$ Resolution (as length) $\times$ Critical Electron Dose]}''~\cite{frank2018single}. 
As these two leaders of the cryo-EM community point out, it is impossible to reconstruct small molecules by any of the existing computational pipelines for single particle analysis in cryo--EM, because the particles themselves cannot be picked from the micrographs. Yet, imaging such molecules is important for a number of applications~\cite{scapin2018cryo}

The unique issues raised by small particles have been mitigated by recent technological advances in the field, including the use of Volta phase plates~\cite{khoshouei2017cryo,liang2017phase} and scaffolding cages~\cite{liu2018nearatomic}.
Despite this progress, detecting small molecules in the micrographs remains a challenge.
We note that nuclear magnetic resonance (NMR) spectroscopy and X-ray crystallography are well suited to reconstruct small molecules. Yet, cryo--EM has a lot to offer even for molecules with already known structures obtained via NMR spectroscopy or X-ray crystallography, because these methods have limited ability to distinguish conformational variability.

@@

As a proof of concept, we first study a toy model  for which it is easier to convey the mathematical principles. 
In this model, an unknown image appears multiple times at unknown locations in each of several micrographs, each affected by additive Gaussian noise---see Figure~\ref{fig:micro_example} for an illustration.
The goal is to estimate the planted image. The task is challenging in particular when the SNR is low enough that particle picking (identifying the locations of each image in each micrograph) cannot be done reliably. 
This problem is interesting on its own as it appears in other scientific applications, including spike sorting~\cite{lewicki1998review}, passive radar~\cite{gogineni2017passive} and system identification~\cite{ljung1998system}.

In order to recover the image, we use autocorrelation analysis. Specifically, we relate the autocorrelations of the micrographs to the autocorrelations of the image.
For any noise level, these autocorrelations can be estimated to any desired accuracy, provided that we  observe sufficiently many image occurrences and the latter are separated in the micrograph. Importantly, there is no need to detect individual image occurrences. The autocorrelations of the micrographs are straightforward to compute and require only one pass over the data. After estimation of the density of particles in the micrographs, these directly yield estimates for the autocorrelations of the target image itself. To estimate the image itself from its estimated autocorrelations, we solve a nonlinear inverse problem; see for instance Figure~\ref{fig:Einst_example}.

Beyond this 2-D proof of concept, we look toward \mbox{3-D} reconstruction as well. Zvi Kam~\cite{kam1980reconstruction} first proposed autocorrelation analysis for \mbox{3-D} reconstruction, under the assumption of perfect particle picking: his method used autocorrelations of the picked, perfectly centered, particles. In contrast, we derive the mathematical relation between the autocorrelations of the micrographs as a whole and the \mbox{3-D} volume, under some simplifying conditions.
We show a few numerical examples and outline the future developments required to make this method applicable to experimental  data.

@@

%\TODO{Need to cite \cite{scapin2018cryo}. Discusses cryo-em for drug discovery and mentions small molecule}

@@

Let $x\in\RL$ be the target signal and let $y\in\RN$ be the observed data, where we assume $N$ is  far larger than $L$. Let  $s \in \{0, 1\}^{N-L+1}$ be a binary signal indicating (with 1's) the starting positions of all occurrences of $x$ in $y$, so that, with additive white Gaussian noise:
\begin{align}
y & =  x \ast s + \varepsilon, & \varepsilon & \sim \mathcal{N}(0,\sigma^2 I_N),
\label{eq:model}
\end{align}
where $\ast$ denotes linear convolution. 
While both $x$ and $s$ are unknown, the goal is only to estimate $x$ from $y$.
This is a key difference with other works on  \emph{blind deconvolution}, a longstanding problem arising in a variety of engineering and scientific applications such as astronomy, communication, image deblurring, system identification and optics; see~\cite{jefferies1993restoration,shalvi1990new,ayers1988iterative,abed1997blind}, just to name a few. 
The parameters of the signal $s$ (the locations of its nonzero values) are the \emph{nuisance variables} of the problem. As will be shown next, in a low SNR environment, estimating $s$ is impossible, whereas estimating $x$ is tractable under some conditions on $s$.

@@

This decision problem can be abstracted as follows: We have two known vectors $\theta_0 = x$ and $\theta_1 = 0$ in $\mathbb{R}^L$. There is a random variable $\eta$ taking values 0 or 1 with probabilities $q$ and $1-q$, respectively. We observe a random vector $X \in \mathbb{R}^L$ (an extract of the micrograph) with the following distribution: if $\eta = 0$, then $X \sim \mathcal N(\theta_0,\sigma^2I_L)$ and if $\eta = 1$, then $X \sim \mathcal N(\theta_1,\sigma^2I_L)$.

@@

In general, for a signal $z$ of length $m$, the autocorrelation of order $q = 1, 2, \ldots$ is given for any integer shifts $\ell_1, \ldots, \ell_{q-1}$ by
\begin{align}
a_z^q[\ell_1,\ldots,\ell_{q-1}]  & = \frac{1}{m} \sum_{i=-\infty}^{+\infty} z[i]z[i+\ell_1]\cdots z[i+\ell_{q-1}],
\label{eq:ac_general}
\end{align}
where indexing of $z$ out of the range $0, \ldots, m-1$ is zero-padded.
For our purposes, this will be applied both to $x$ (of length $L$) and to $y$ (of length $N$).
Explicitly, the first-, second- and third-order autocorrelations are given by: %\TODO{we may want to remove these explicit formulas}
\begin{align} 
a_z^1 & = \frac{1}{m} \sum_{i=0}^{m-1} z[i], \nonumber\\
a_z^2[\ell] & = \frac{1}{m} \sum_{i = \max\{0, -\ell\}}^{m-1 + \min\{0, -\ell\}} z[i]z[i+\ell], \nonumber\\
a_z^3[\ell_1,\ell_2] & = \frac{1}{m} \sum_{i = \max\{0, -\ell_1, -\ell_2\}}^{m-1 + \min\{0, -\ell_1, -\ell_2\}} z[i]z[i+\ell_1]z[i+\ell_2]. \label{eq:ac_special}
\end{align}
The autocorrelation functions have symmetries. Specifically, $a_z^2[\ell] = a_z^2[-\ell]$, and
$a_z^3[\ell_1,\ell_2] = a_z^3[\ell_2,\ell_1]=a_z^3[-\ell_1,\ell_2-\ell_1].
$

@@

For shifts in $0, \ldots, L-1$, the autocorrelation functions of $y$ depend on the corresponding autocorrelations of $x$, the noise level $\sigma$ and the support signal $s$. Importantly, under the separation condition~\eqref{eq:spacing}, the dependency on $s$ is only through the density $\gamma$.
We consider the asymptotic regime where $\gamma$ remains constant; that is, as $N$ goes to infinity, $M$ also goes to infinity at the same rate (in other words, as we see an increasingly large micrograph, it contains increasingly many signal occurrences, with constant signal density). In that regime, the law of large numbers can be used to show the following statement:
\begin{align} \label{eq:mean_micrograph}
\lim_{N\to\infty} a_y^1 & \aseq \gamma a_{x}^1,
\end{align}
where equality holds almost surely (a.s.), meaning it holds with probability one.
The randomness is over the Gaussian noise $\varepsilon$; $s$ may be deterministic.
Thus, given enough data, if $\gamma$ is known, we can estimate $a_x^1$ from $y$. (We show later how to estimate $\gamma$ as well.)

We have a similar observation for the second-order autocorrelation: $a_y^2[\ell]$ computes the correlation between $y$ and a copy of $y$ shifted by $\ell$ entries. Considering $\ell$ only in the range $0, \ldots, L-1$, one can see that any given occurrence of $x$ in $y$ is only ever correlated with itself, and never with another occurrence. As a result,
\begin{align}
	\lim_{N\to\infty} a_y^2[\ell] & \aseq \gamma a_{x}^2[\ell] + \sigma^2\delta[\ell],
	\label{eq:ac2_micrograph}
\end{align}
for $\ell = 0, \ldots, L-1$, where $\delta[\ell]$ equals one for $\ell=0$ and zero otherwise. The last part captures the autocorrelation of the noise. Notice that, even if $\sigma$ is unknown, entries $\ell = 1, \ldots, L-1$ still provide useful information about $a_x^2$.

Along the same lines, one can establish a relation for third-order autocorrelations:
\begin{multline}
\lim_{N\to\infty} a_y^3[\ell_1,\ell_2] \aseq \gamma a_{x}^3[\ell_1,\ell_2] \label{eq:ac3_micrograph} \\ + \sigma^2\gamma a_{x}^1 \cdot \big(\delta[\ell_1,0]+\delta[0,\ell_2]+\delta[\ell_1,\ell_2]\big),
\end{multline}
for $\ell_1,\ell_2 = 0, \ldots, L-1$, where $\delta[\ell_1,\ell_2]:=\delta[\ell_1-\ell_2]$.
Here too, few entries are affected by $\sigma$ in the limit.
See Appendix~\ref{sec:autocorrelation_computation}.

@@

A couple of remarks are in order. First, \eqref{eq-uniqueness} is not numerically stable: if $z[0]$ or $z[L-1]$ are close to 0, recovery of $z$ is sensitive to errors in the autocorrelations. In practice, we recover $z$ by fitting it to its autocorrelations using a nonconvex least-squares (LS) procedure, which is empirically more robust to additive noise; we have observed similar phenomena for related problems~\cite{bendory2017bispectrum,boumal2017heterogeneous,abbe2017multireference}.
Second, note that the second-order autocorrelation is not by itself sufficient to determine the signal uniquely. %~\cite{beinert2015ambiguities,bendory2017fourier}.
However, for dimensions greater than one, almost all signals are determined uniquely up to sign (or phase for complex signals) and reflection through the origin (with conjugation in the complex case)~\cite{hayes1982reconstruction}. %,hayes1982reducible}. 
The sign ambiguity can be resolved by the mean of the signal if it is not zero. However, determining the reflection symmetry still requires additional information, beyond the second-order autocorrelation.

The observed moments $a_y^1,a_y^2$ and $a_y^3$ of $y$ do not immediately yield the moments of the signal $x$, as seen by~\eqref{eq:mean_micrograph},~\eqref{eq:ac2_micrograph} and~\eqref{eq:ac3_micrograph}; rather, the two are related by the noise level $\sigma$ and the ratio $\gamma$. We will show, however, that $x$ is still determined by the observed moments of $y$.
%
%identifiable from the observed moments of $y$. In general, we say a parameter is ``identifiable'' if its value is uniquely determined in the limit $N \to \infty$.

First, we observe that if the noise level $\sigma$ is known, generally, one can estimate $\gamma$ from the first two moments of the micrograph.
%
\begin{proposition} \label{prop:gamma}
	Let $\sigma > 0$ be fixed and assume that the separation condition~\eqref{eq:spacing} holds. If the mean of $x$ is nonzero, then 
	%
	\begin{align*}
	%
	\gamma & \aseq \lim_{N \to \infty}\frac{L (a^1_y)^2}{a_y^2[0] + 2\sum_{\ell = 1}^{L-1}a_y^2[\ell]-\sigma^2}.
	%
	\end{align*}
	%
\end{proposition}
\begin{proof}
In the limit, $(a^1_y)^2=\frac{\gamma^2}{L^2}\sum_{i=0}^{L-1}\sum_{j=0}^{L-1}x[i]x[j]$. Similarly,  $\sum_{\ell = 1}^{L-1}a_y^2[\ell]=\frac{\gamma}{L}\sum_{\ell = 1}^{L-1}\sum_{i=0}^{L-1-\ell}x[i]x[i+\ell]$ and $a_y^2[0]=\frac{\gamma}{L}\sum_{i=0}^{L-1}x^2[i] + \sigma^2$. The proof is concluded by noting that  $a_x^2[-\ell]=a_x^2[\ell]$. 
%	Plug in the explicit expressions of~\eqref{eq:mean_micrograph} and~\eqref{eq:ac2_micrograph} into the right hand side of the equality.
\end{proof}

Using third-order autocorrelation information of $y$, both the ratio $\gamma$ and the noise $\sigma$ are determined. For the following results, when we say that a result holds for a ``generic'' signal $x$, we mean that the set of signals which cannot be determined by these measurements
has Lebesgue measure zero. 
In particular, this means that we can recover
almost all signals with the given measurements.

\begin{proposition} \label{prop:gamma_sigma}
	%
	%Let $\sigma > 0$ be fixed. 
	Assume $L \geq 3$ and assume that the separation condition~\eqref{eq:spacing} holds. % Added by Nicolas, Sep 4, 2018: I checked all identities for L >= 3, with some computer code; I didn't write code to verify the corner cases L = 1 and L = 2.
	In the limit of $N\to \infty$, the observed autocorrelations $a_y^1,a_y^2$ and  $a_y^3$ determine the ratio $\gamma$ and noise level $\sigma$ uniquely for a generic signal $x$. If $\gamma > \frac{1}{4}$, then this holds for any signal $x$ with nonzero mean. 
	\begin{proof}
		See Appendix~\ref{sec:proof_prop_gamma_sigma}.
	\end{proof}
\end{proposition}

From Propositions~\ref{prop:uniqueness} and~\ref{prop:gamma_sigma} we   deduce the following:
\begin{corollary}
	%Let $K=1$ and $\sigma > 0$ be fixed.
	 In the limit of $N\to \infty$ and under the separation condition~\eqref{eq:spacing}, the signal $x$, the ratio $\gamma$, and the noise level $\sigma$ are determined from the first three autocorrelation functions of $y$ if:
	\begin{itemize}
		\item Either the signal $x$ is generic; or
		\item   $x$ has nonzero mean  and $\gamma > \frac{1}{4}$.
	\end{itemize}
\end{corollary}
%
%\TODO{Similar results hold for signals in higher dimensions.---need to keep this sentence in red until the very end to make sure it's correct.}

As a side note, under the separation condition, the length $L$ of the signal can also be determined from the autocorrelations in the asymptotic regime, by inspection of the support of $a_y^2$.
%In particular, if  $a_y^2[L-i] = 0$ for $i=1,\ldots,T$,
%then the support of $x$ has length $L-T$ rather than $L$.

@@

\section*{Acknowledgment}
The authors  thank Ayelet Heimowitz, Joe Kileel,  Roy Lederman, Amit Moscovich and Nir Sharon for helpful discussions, and Fred Sigworth for supplying us with the dataset of the BPTI molecule. The research was partially supported by Award Number R01GM090200 from the NIGMS, FA9550-17-1-0291 from AFOSR, Simons Foundation Math+X Investigator Award, and the Moore Foundation Data-Driven Discovery Investigator Award.
\revised{Amit: to verify.} NB is partially supported by NSF award DMS-1719558.
\TODO{Correct bib}

@@

We run the toy model experiments on a shared computer with 144 logical CPUs of type Intel(R) Xeon(R) CPU E7-8880 v3 @ 2.30GHz and 792 Gb of RAM; we use at most 72 of these CPUs.
